Phylogenetics and the human microbiome.

### 2013-09-25 sbreview.tex
Introduction roughed out.

Why, as systematic biologists, should we be interested in what these microbial ecologists have done?
System with a lot of data and a lot of questions.
Microbes can only be identified by their molecular sequence.
Indeed, just to do ecology you have to do sequencing.
Many of the issues that are raised by high throughput sequencing of beasties that are more commonly of interest for phylogenetics folk are present, and worse, in microbial studies.

Taxonomic issues.
Lack of "obvious" ways of grouping beasties.
Lack of broad phenotypic characterization.
In the absence of sequencing, microbiologist have to rely on microscopes and staining.
Lots of little things look alike.
These issues motivate sequence- and evolution-based methodology.

This has led to innovation in ways of using trees.
It means something that they have developed their own way of using trees, although there is some overlap with previous literature.
Questions of microbial genomics, translated into questions of function, have important consequences for human health.

Why human microbiome versus microbial ecology?
Relevance to health.
We know a lot more about these bugs.
Can manipulate the systems-- feeding studies and gnotobiotic mice.

In this review I will describe human microbiome research from the perspective of a phylogenetics researcher.
I will first briefly review the recent literature on the human microbiome, then describe novel ways in which human microbiome researchers have used trees and the challenges that human microbiome research poses for the phylogenetics community.
I will finish with opportunities for SB audience to contribute to this field.

@@

\section{Methods used to investigate the human microbiome via sequencing.}

Terminology: survey versus metagenome.
Many of these methods cross over with microbial ecology in general \citep{hugenholtz2002review} (cite this?).
I will not be talking about manipulations of these systems, such as gnotobiotic mice and human intervention.

@@

@@

While Woese and colleagues used primitive sequencing methods to gain information, modern researchers have the luxury of high throughput sequencing.
This can be done with a high level of multiplexing, making an explicit trade-off between depth of sequencing for each specimen and the number of specimens able to be put on the sequencer at the same time.
This leads to the question of how many sequences are needed to characterize the microbial diversity of a given environment.
For simply separating two rather different samples, rather few samples are needed \cite{kuczynski2010microbial}.

>>>>

The parameter regime and focus of microbiome research sits outside of the traditional setting for systematic biology, thus why should our community be interested in what these microbial ecologists and medical researchers have done?
This system is data- and question-rich.
It absolutely requires molecular methods, as microbes can be identified by their molecular sequence, which is much more straightforward to do in high throughput than morphological or phenotypic characterization.
Indeed, just to do ecology you have to do sequencing.
Many of the issues that are raised by high throughput sequencing of beasties that are more commonly of interest for phylogenetics folk are present, and worse, in microbial studies.
For instance, the diversity of these organisms is great and organizing them into a taxonomy is a serious challenge, especially in the absence of obvious morphological features.

Because of challenges with species definitions, microbial ecology researchers have developed their own techniques for doing comparing samples which are explicitly phylogenetic.
Although there is some overlap with previous literature, these techniques are novel and could be used in a wider setting and deserve wider consideration by the phylogenetics community.

The human microbiome part of microbial ecology is specifically interesting because questions of microbial genomics, translated into questions of function, have important consequences for human health.
Additionally, due to more than a century of hospital lab medicine, our knowledge about human-associated microbes is much richer than about microbes in general.
It is relatively routine to manipulate the human microbiome or models thereof via intervention studies and germ-free animals.

In this review I will describe human microbiome research from the perspective of a phylogenetics researcher.
I will first briefly review the recent literature on the human microbiome, then describe novel ways in which human microbiome researchers have used trees and the challenges that human microbiome research poses for the phylogenetics community.
I will finish with opportunities for the \textit{Systematic Biology} audience to contribute to this field.

@@

\section{Investigating the human microbiome via sequencing}

Terminology: survey versus metagenome.
Many of these methods cross over with microbial ecology in general \citep{hugenholtz2002review} (cite this?).
% I will not be talking about manipulations of these systems, such as gnotobiotic mice and human intervention.

@@

Definition. Goal.

@@

While Woese and colleagues used primitive sequencing methods to gain information, modern researchers have the luxury of high throughput sequencing.
This can be done with a high level of multiplexing, making an explicit trade-off between depth of sequencing for each specimen and the number of specimens able to be put on the sequencer at the same time.
This leads to the question of how many sequences are needed to characterize the microbial diversity of a given environment.
For simply separating two rather different samples, rather few samples are needed \cite{kuczynski2010microbial}.
This has led to desire for massive parallelization (cite).
